\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-04-18 paper/main.tex
.

\abstract{\textbf{Motivation:} \textit{In silico} drug--target
  interaction (DTI) prediction is important for drug discovery and
  drug repurposing.  Approaches to predict DTIs can proceed
  indirectly, top-down, using phenotypic effects of drugs to identify
  potential drug targets, or they can be direct, bottom-up and use
  molecular information to directly predict binding potentials.
  Both approaches can be combined with information about interaction
  networks,

  However, DTI datasets are prone to certain biases,
  skewing the effective predictive expressiveness over those
  datasets.
  \textbf{Results:} We present \name{} to combine molecular features
  and functional information with protein--protein interaction
  networks under usage of graph convolutional neural networks, showing
  that phenotypic information localizes on the interaction network. We
  hereby focus on protein-centric evaluation, predicting drugs that
  may target specific targets. We further show that common
  drug--target interaction datasets contain intrinsic biases, majorly
  affecting predictive performance of compared models. Additionally,
  we propose an modified evaluation scheme and metric to circumvent
  such skews and to better reflect the problem DTI prediction methods
  aim to solve. \\
  \textbf{Availability:} Source code and data used in \name{} are
  available at \url{https://github.com/THinnerichs/DTI-VOODOO} \\
  \textbf{Contact:}
  \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}
  \\
  \textbf{Supplementary information:}Supplementary data are
  available at \textit{Bioinformatics} online.}

>>>>

\abstract{\textbf{Motivation:} \textit{In silico} drug--target
  interaction (DTI) prediction is important for drug discovery and
  drug repurposing.  Approaches to predict DTIs can proceed
  indirectly, top-down, using phenotypic effects of drugs to identify
  potential drug targets, or they can be direct, bottom-up and use
  molecular information to directly predict binding potentials.  Both
  approaches can be combined with information about interaction
  networks. \\
  \textbf{Results:} We developed \name{} as a computational method
  that combines molecular features and ontology-encoded phenotypic
  effects of drugs with protein--protein interaction networks, and
  uses a graph neural networks to predict DTIs.  We demonstrate that
  drug effect features can exploit information in the interaction
  network whereas molecular features do not.  \name{} is designed to
  predict candidate drugs for a given protein; we use this formulation
  to show that common DTI datasets contain intrinsic biases with major
  affects on performance evaluation and comparison of DTI prediction
  methods. Using a modified evaluation scheme, we demonstrate that
  \name{} improves substantially over state of the art DTI prediction
  methods. \\
  \textbf{Availability:} \name{} source code and data necessary to
  reproduce results are freely available at
  \url{https://github.com/THinnerichs/DTI-VOODOO}. \\
  \textbf{Contact:}
  \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}
  \\
  \textbf{Supplementary information:}Supplementary data are available
  at \textit{Bioinformatics} online.}
